Highly Available Transactions: Virtues and Limitations

### 2013-03-31 writeup/intro.tex
typos

Indeed, serializable transactions, the gold standard of traditional
ACID databases, is indeed not achievable given a requirement for high
availability in the presence of
partitions~\cite{davidson-survey}. However, database systems have a
long tradition of providing weaker isolation and consistency
guarantees~\cite{adya, ansicritique, gray-isolation,
  gray-virtues} (TODO: kemme). Due to their concurrency and overall performance
benefits, weak isolation models are widely employed, overwhelmingly as
the default setting in today's traditional ``ACID'' and up-and-coming
``NewSQL'' databases and often as the maximum guarantee offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics of current highly available systems. This raises a natural
question: which of these alaternative semantics can be provided while
remaining highly available?

@@

More precisely, the reason why fairly strong models like Snapshot
Isolation (SI) and Repeatable Reads (RR) are not highly available is
that they require detecting conflicts between concurrent updates (as
needed for preventing Lost Updates or Write Skew phenomena). However,
Read Committed isolation, Transactional Atomicity, and many of the
weaker data consistency models from database and distributed systems
can be implemented in a highly available manner. For several of these
guarantees, such as Transactional Causal Predicate Cut Isolation, we
consider a modified form of high availiabity in which clients ``stick
to'' (i.e., have affinity with) at least one server (a property which
is often implicit in the distributed systems
literature~\cite{herlihy-art, cops, eiger, cac} but which requires
\textit{explicit} consideration in a client-server replicated database
context). This sticky availability is easily implementatable and
widely employed~\cite{cops, eiger, vogels-defs} but is strictly less
available than traditional definitions of high availability.

>>>>

Indeed, serializable transactions, the gold standard of traditional
ACID databases, is indeed not achievable given a requirement for high
availability in the presence of
partitions~\cite{davidson-survey}. However, database systems have a
long tradition of providing weaker isolation and consistency
guarantees~\cite{adya, ansicritique, gray-isolation,
  gray-virtues} (TODO: kemme). Due to their concurrency and overall performance
benefits, weak isolation models are widely employed, overwhelmingly as
the default setting in today's traditional ``ACID'' and up-and-coming
``NewSQL'' databases and often as the maximum guarantee offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics of current highly available systems. This raises a natural
question: which of these alternative semantics can be provided while
remaining highly available?

@@

More precisely, the reason why fairly strong models like Snapshot
Isolation (SI) and Repeatable Reads (RR) are not highly available is
that they require detecting conflicts between concurrent updates (as
needed for preventing Lost Updates or Write Skew phenomena). However,
Read Committed isolation, Transactional Atomicity, and many of the
weaker data consistency models from database and distributed systems
can be implemented in a highly available manner. For several of these
guarantees, such as Transactional Causal Predicate Cut Isolation, we
consider a modified form of high availability in which clients ``stick
to'' (i.e., have affinity with) at least one server (a property which
is often implicit in the distributed systems
literature~\cite{herlihy-art, cops, eiger, cac} but which requires
\textit{explicit} consideration in a client-server replicated database
context). This sticky availability is easily implementable and
widely employed~\cite{cops, eiger, vogels-defs} but is strictly less
available than traditional definitions of high availability.
